
# Investment evaluation as measure of forecast success

We want to see whether certain forecasts perform better than random ones in making an investment decision on stocks. 

First we define 2 functions which invest when we predict a rise and calculate the winnings.

```{r}
library(forecast)

invest_eval <- function(last, pred, actual) {
  # if forecast > last invest, calc the return (sell next period)
  if (pred > last) {
  	return(actual - last)
  } else {
  	return(0)
  }
}

wins <- function(x, p, idx) {
  stopifnot(max(idx) <= length(x))
  w <- sapply(idx, function(i) {
    invest_eval(x[i - 1], p[i], x[i])
  })
  return(w)
}

set.seed(100)
n <- 100

# 2 uncorrelated random series
xrand <- rnorm(n, 10, 1)
prand <- rnorm(n, 10, 1)
cor(xrand, prand)

# 2 uncorrelated trended series
xtrend <- 1:n / 5 + rnorm(n, 0, 1)
ptrend <- 1:n / 5 + rnorm(n, 0, 1)
cor(xtrend, ptrend)

# 2 process with same structure but not related
xar1 <- 10 + arima.sim(list(ar = .8), n)
par1 <- 10 + arima.sim(list(ar = .8), n)
cor(xar1, par1)

# auto.arima forecast
par1f <- sapply(11:n, function(i) {
  # 1:10 -> 11
  forecast(auto.arima(xar1[1:(i - 1)]), 1)$mean
})
par1f <- c(rep(NA, 10), par1f)

# theta forecast
par1f2 <- sapply(11:n, function(i) {
  # 1:10 -> 11
  thetaf(xar1[1:(i - 1)], 1)$mean
})
par1f2 <- c(rep(NA, 10), par1f2)

# naive oppose the sign 
par1f3 <- sapply(11:n, function(i) {
  # oppose the last change
  xar1[i - 1] - sign(xar1[i - 1] - xar1[i - 2])
})
par1f3 <- c(rep(NA, 10), par1f3)

cor(xar1, par1, use = 'compl')
cor(xar1, par1f, use = 'compl')
cor(xar1, par1f2, use = 'compl')
cor(xar1, par1f3, use = 'compl')

# calc wins
wrand <- wins(xrand, prand, 11:n)
wtrend <- wins(xtrend, ptrend, 11:n)
war1 <- wins(xar1, par1, 11:n)
war1f <- wins(xar1, par1f, 11:n)
war1f2 <- wins(xar1, par1f2, 11:n)
war1f3 <- wins(xar1, par1f3, 11:n)

matplot(cbind(xrand, prand), type = 'l', main = 'random')
matplot(cbind(xtrend, ptrend), type = 'l', main = 'trend')
matplot(cbind(xar1, par1, par1f, par1f2, par1f3), type = 'l', main = 'AR1')
legend('bottomleft', c('ar1', 'random', 'auto.arima', 'thetaf', 'opp sign'), lty = 1, col = 1:5)

# plot cum wins
plot(cumsum(wrand), type = 'l')
plot(cumsum(wtrend), type = 'l')
matplot(cbind(cumsum(war1), cumsum(war1f), cumsum(war1f2), cumsum(war1f3)), type = 'l')
legend('bottomleft', c('random', 'auto.arima', 'thetaf', 'opp sign'), lty = 1, col = 1:4)

# sum wins
sum(wrand)
sum(wtrend)
sum(war1)
sum(war1f)
sum(war1f2)
sum(war1f3)

# what is structure of random and trended series
summary(auto.arima(xrand))
summary(auto.arima(xtrend))
```

Note that all series have mean of 10, so wins are kind of comparable. Using uncorrelated data for random and trended data, gives positive returns.

For an AR(1) the random process gives the best returns, 3 forecast methods give worse returns. 

```{r}
# now try with real data

library(quantmod)
set.seed(100)
getSymbols('DIS')
plot(DIS)
x2 <- as.data.frame(DIS)[1:n, 4]

summary(auto.arima(x2))

# a unrelated process with same structure
parma11 <- mean(x2) + arima.sim(list(order = c(1, 1, 1), ar = .59, ma = -.84), n, sd = sd(x2))[1:n]
matplot(cbind(x2, parma11), type = 'l')
legend('bottomleft', c('DIS', 'ARIMA(1,1,1)'), lty = 1, col = 1:2)

px2f <- sapply(11:n, function(i) {
  # 1:10 -> 11
  forecast(auto.arima(x2[1:(i - 1)]), 1)$mean
})
px2f <- c(rep(NA, 10), px2f)

px2f2 <- sapply(11:n, function(i) {
  # 1:10 -> 11
  thetaf(x2[1:(i - 1)], 1)$mean
})
px2f2 <- c(rep(NA, 10), px2f2)

px2f3 <- sapply(11:n, function(i) {
  # oppose the last change
  x2[i - 1] - sign(x2[i - 1] - x2[i - 2])
})
px2f3 <- c(rep(NA, 10), px2f3)

cor(x2, parma11, use = 'compl')
cor(x2, px2f, use = 'compl')
cor(x2, px2f2, use = 'compl')
cor(x2, px2f3, use = 'compl')

matplot(cbind(x2, parma11, px2f, px2f2, px2f3), type = 'l')
legend('bottomleft', c('actual', 'random', 'auto.arima', 'thetaf', 'opp sign'), lty = 1, col = 1:5)

wx2 <- wins(x2, parma11, 11:n)
wx2f <- wins(x2, px2f, 11:n)
wx2f2 <- wins(x2, px2f2, 11:n)
wx2f3 <- wins(x2, px2f3, 11:n)

matplot(cbind(cumsum(wx2), cumsum(wx2f), cumsum(wx2f2), cumsum(wx2f3)), type = 'l')
legend('bottomleft', c('random', 'auto.arima', 'thetaf', 'opp sign'), lty = 1, col = 1:4)
sum(wx2)
sum(wx2f)
sum(wx2f2)
sum(wx2f3)
```

The random process and the Arima forecast have positive returns. Sign forecast has negative returns. 
